What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-13 android-stacks.tex
typpos

This paper reports a study which mined the Java stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). The goal of this study is to investigate 
whether the analysis of stack trace information could reveal bug hazards related to the exception handling code,
 that may lead to a decrease in application robustness. 
 Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some of the outcomes of this study include the identification of the following 
bug hazards: (i) unexpected exception wrappings (such as Errors being wrapped in checked exceptions) -
indicating that Java's hybrid exception model become fault-prone when it is not fully used in 
accordance with its purpose; (ii) undocumented runtime exceptions thrown by both the Android 
platform and third party libraries; and (iii) undocumented checked exceptions thrown by the Android Platform.
This study provides insights on such bug hazards and the robustness threats  
they impose not only to Android apps but also to other systems based on Java exception model.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions.
In this paper we will be concerned with exception handling in Android apps,
which reuse Java's exception handling model.

@@

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
the thrown exception and the ordered list of methods that were active on the call stack before 
such exception has occurred ~\cite{gosling2000java,bloch2008effective}.

@@

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common bug hazards in the exception-related code.  
A bug hazard  ~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug to be present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.

To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of bytecode 
and source code analysis on the exception-related code of the Android 
platform and Android applications. Some bug hazards consistently detected 
during this mining study include:

@@

   \item  Cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception.
Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
 to an unpredictable state. Such wrappings suggest that, when not adequately used, the Java's hybrid exception
 model may negatively impact the application robustness.

@@

   \item Undocumented checked exceptions signaled by native C code.  Some flows reported a checked 
exception signaled by native C code invoked by the Android Platform, yet this exception was not declared
 of the Java Native Interface. This can lead to uncaught exceptions that are 
difficult to debug.

@@

 The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, some of the findings of our
 study emphasize the impact of these bug harzards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while other works analyzed crash and automated test reports.
 On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.

Such outcomes point to threats not only to the development of robust Android apps, 
but also to the development of any robust Java-based system. 
Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
 bug hazards the application robustness, and how tricky it can be to prevent them. Moreover, such bug hazards 
calls for improvements on languages and tools to deal with such bug hazards in Android and Java environments.

@@

\subsection{The Android Platform} \label{sec:extypes}
Android is an open source platform for mobile devices based on Linux kernel,
which also comprises a subset of the core Java library, a set of third-party libraries (e.g.  SQLite),
the Dalvik runtime environment, and the Application Framework (which provides 
higher-level services to the applications running on the platform) ~\cite{andguide}.
The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask) - 
which are usually needed due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

@@

\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, on which
 every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. The checked exception received such name
 because they must be declared on the method's exception interface (i.e., a list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as unchecked exceptions, because 
they do not need to be specified on method exception interface and do not trigger any 
compile time checking.

By convention, instances of Error represent  unrecoverable conditions which usually results
from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
normally they cannot be handled inside the application.  Instances of RuntimException are implicitly 
thrown by Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. 
The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
to the Java Specification ~\cite{gosling2000java}. The user-defined exceptions can be either checked
 or runtime. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section ~\ref{sec:best} presents
a set of best practices related to each exception type.

@@

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates an exception wrapping. 
For simplicity we will refer to exception stack trace as stack trace along the text.
The bottom part of the stack trace is the \emph{root exception}, which indicates
the first reason (root cause) for the exception be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{exception wrapper} along this paper). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. Such guidelines do not focus on 
advocating any specific exception type, but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

%\textbf{I
\emph{I-Checked exceptions should be used to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

@@

\emph{IV- All exceptions explicitly thrown by reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document (either or on the exception interface or using @throws tag on Javadoc) 
the explicitly thrown runtime exceptions as carefully as checked exceptions. 
When mentioning reffering to `` explicit" thrown exceptions
we  exclude the implicitly signaled runtime exceptions related to programming
mistakes. Such practice is mostly needed when developing library/framework 
code reusable code - as the developers do not have access to the source code. 
Doing so, the clients of such library/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

If the developer fails to follow such practices (specially when developing 
library/framework code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code,
which can lead to the uncaught exceptions - one of the main causes of 
systems crashes. These best practices guided our exploratory study described next.

@@

%This section describes the exploratory study whose goal was to investigate:
%\begin{description}
 % \item[RQ] \noindent\emph{Which fragilities on the exception-related code can be revealed from a post morten exception stack trace analysis of Android projects?} 
 %    \item
%\end{description}
%One major decision that had to be made for our investigation was the selection of the target 
%projects. We have selected a set of XXX open-source Android projects hosted in GitHub and 
%Google Code. GitHub and Google Code are among the most used hosting sites for open-source
 %projects,

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
This exploration is guided by the set of best practices compiled in this study. We investigate whether characteristics 
of exception stack traces can pinpoint best practices violations and therefore reveal bug hazards.
As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
that favor the introduction of failures such as uncaught exceptions.

To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.

Figure~\ref{fig:overview} gives an overview of the study.
Firstly  the issues reported on Android projects hosted on GitHub and Google Code were recovered (Figure~\ref{fig:overview} - (1) and (2)).
Then the stack traces embedded each issue were extracted and distilled (Figure~\ref{fig:overview} - (3)).
The stack trace information were then combined with source code and bytecode analysis in order to discover 
the type of the exceptions reported on the stack traces (e.g., error, runtime, checked)  (Figure~\ref{fig:overview} - (5))  and the origin of such exceptions 
(e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
Manual inspections were also used to support this mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and support
the search for bug hazards performed in this study  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool 
 in this set we observed that 589 projects had at least one issue containing a stack trace.

@@

Based on the assumption that regardless the issue lables, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect issues.

@@

The Crawler retrieved all issues and comments for these projects.
From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further cleanup (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stacks (regardless its labels).

@@

The ExceptionMiner is a mining tool which can connect to different issue repositories, 
extract issues, mine stack traces from them, distill exception stack trace information,
and enable the execution of different analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
 that can traverse Google Code Web interface and extract the project issues. 
Projects meta-data and the issues associated to each project are stored on a relational 
database.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces mixed with log information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.}.

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analyze the source code of Java Exception classes (when the
bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed all exceptions defined on Android platform (API level 19) - 
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover it also analyzed the 
libraries and applications which defined exceptions reported on stack traces
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could refer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, the tool recursively analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source code of reused libraries.
Then based on this information and using pattern matching 
between the signaler name and the packages this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were not defined 
on Android platform, or on core libraries or on applications.
Table~\ref{tab:signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used to:
(i) support on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and (ii)  
identify the type of some exceptions reported on issues, 
that were not be automatically identified by the ExceptionMiner tool
(because they were defined on previous versions of libraries,
apps and Android Plaform). When the exception could not be  
found niether automatically, nor manually (because they were defined on a previous version
of the app or lib) the exception were classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

\subsection{Replication Package}
All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guided the search for bug hazards on the exception-related code
based on exception stack trace information.

@@

After distilling the information available on the exception stack traces we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

>>>>

This paper reports a study which mined the exception stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). The goal of this study is to investigate 
whether the analysis of stack trace information could reveal \emph{bug hazards} related to the exception handling code,
 that may lead to a decrease in application robustness. 
 Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some of the outcomes of this study include the identification of the following 
\emph{bug hazards}: (i) unexpected cross-type exception wrappings (for instance, trying to handle an instance 
of OutOfMemoryError ``hidden" in  a checked exception)  which can make the exception-related code 
more complex and negatively impact the application robustness; (ii) undocumented runtime exceptions 
thrown by both the Android platform and third party libraries; and (iii) undocumented checked exceptions
 thrown by the Android Platform. Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen while calling a third-party code.
This study provides further insights on such \emph{bug hazards} and the robustness threats  
they impose not only to Android apps but also to other systems based on Java exception model.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions.
In this paper we will be concerned with exception handling in Android apps,
which reuses Java's exception handling model.

@@

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or on a log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
of the thrown exception and the ordered list of methods that were active on the call stack before 
such exception has occurred ~\cite{gosling2000java,bloch2008effective}.

@@

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common \emph{bug hazards} in the exception-related code.  
A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug to be present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.

To guide this investigation we compiled general guidelines on how to use Java
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of bytecode 
and source code analysis on the exception-related code of the Android 
platform and Android applications. Some \emph{bug hazards} consistently detected 
during this mining study include:

@@

   \item  Cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception.
Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
 to an unpredictable state. Such wrappings suggest that when (mis)applied, the exception wrapping can make 
the exception-related code more complex and negatively impact the application robustness.

@@

   \item Undocumented checked exceptions signaled by native C code.  Some flows reported a checked 
exception signaled by native C code invoked by the Android Platform, yet this exception was not declared
 of the Java Native Interface invoking it. This can lead to uncaught exceptions that are 
difficult to debug.

@@

 The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, some of the findings of our
 study emphasize the impact of these bug harzards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while the previous works analyzed crash reports and automated test reports.
 On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to the application robustness.

Such outcomes point to threats not only to the development of robust Android apps, 
but also to the development of any robust Java-based system. 
Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
 \emph{bug hazards} the application robustness, and how tricky it can be to prevent them. Moreover, such \emph{bug hazards} 
calls for improvements on languages and tools to better deal with them in Android and Java environments.

@@

\subsection{The Android Platform} \label{sec:extypes}
Android is an open source platform for mobile devices based on Linux kernel,
which also comprises (i) a set of native libraries written in C/C++ 
(e.g., WebKit, OpenGL, FreeType, SQLite, Media, C runtime library) - which 
fulfill a wide range of functions including graphics drawing, SSL communication, 
SQLite database management, audio and video playback etc; (ii) a set of Java Core Libraries 
- which include a subset of Java standard libraries and mainly “wrappers” to access the set of C/C++ 
native libraries using the Java Native Interface (JNI); (iii) the Dalvik runtime environment - 
which was specifically designed to deal with the resource constraints of a mobile device; 
 and (iv) the Application Framework which provides higher-level APIs to the applications
 running on the platform.

@@

\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, on which
 every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. The checked exception received such name
 because they must be declared on the method's \emph{exception interface} (i.e., a list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as unchecked exceptions, because 
they do not need to be specified on method \emph{exception interface} and do not trigger any 
compile time checking.

By convention, instances of Error represent  unrecoverable conditions which usually result
from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
normally they cannot be handled inside the application.  Instances of RuntimException are implicitly 
thrown by Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g., C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. 
According to the Java Specification~\cite{gosling2000java} the programs are not 
expected to handle such runtime exceptions signaled by the runtime environment. 
The user-defined exceptions can be either checked
 or runtime. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section ~\ref{sec:best} presents
a set of best practices related to each of them.

@@

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} shows 
an exception stack trace which illustrates an exception wrapping. 
For simplicity along the text we will refer to exception stack trace as stack trace.
The bottom part of the stack trace is the \emph{root exception}, which indicates
the first reason (root cause) for the exception be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{exception wrapper} along this paper). The
execution flow  between the root exception and the wrapper may
include other intermediate exception wrappers. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. Such guidelines do not focus on 
advocating any specific exception type, but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below\footnote{We also compiled aditional guidelines related
to ways of handling exceptions but they are out of the scope of this paper.}:

@@

%\textbf{I
\emph{I-Checked exceptions should be used to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

@@

\emph{IV- All exceptions explicitly thrown by reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document (either on the \emph{exception interface} or using @throws tag on Javadoc) 
the \emph{explicitly thrown} runtime exceptions as carefully as checked exceptions. 
The \emph{explicitly thrown} exceptions exclude the ones implicitly 
thrown by the runtime environment due to programming
mistakes. Such practice is mostly needed when developing library/framework 
code reusable code - as the developers do not have access to the source code. 
Doing so, the clients of such library/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}. If the developer fails to follow this practice 
 it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code. 
Such exceptions can lead to the uncaught exceptions - one of the main causes of 
systems crashes.

@@

%This section describes the exploratory study whose goal was to investigate:
%\begin{description}
 % \item[RQ] \noindent\emph{Which fragilities on the exception-related code can be revealed from a post morten exception stack trace analysis of Android projects?} 
 %    \item
%\end{description}
%One major decision that had to be made for our investigation was the selection of the target 
%projects. We have selected a set of XXX open-source Android projects hosted in GitHub and 
%Google Code. GitHub and Google Code are among the most used hosting sites for open-source
 %projects,
% We investigate whether characteristics 
%of exception stack traces can pinpoint best practices violations and therefore reveal \emph{bug hazards}.

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
As mentioned before, in this context \emph{bug hazards} are the characteristics on the exception-related code 
that favor the introduction of failures such as uncaught exceptions.

This exploration is guided by the set of best practices compiled in this study.
To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).

Figure~\ref{fig:overview} gives an overview of the study.
Firstly  the issues reported on Android projects hosted on GitHub and Google Code were recovered (Figure~\ref{fig:overview} - (1) and (2)).
Then the stack traces embedded each issue were extracted and distilled (Figure~\ref{fig:overview} - (3)).
The stack trace information were then combined with source code and bytecode analysis in order to discover 
the type of the exceptions reported on the stack traces (e.g., error, runtime, checked)  (Figure~\ref{fig:overview} - (5))  and the origin of such exceptions 
(e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
Manual inspection steps were used to support the mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and 
the search for \emph{bug hazards}  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term ``android" in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool 
 in this set we observed that 589 projects had at least one issue containing a stack trace.

@@

Based on the assumption that regardless the issue lables, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review \emph{bug hazards} on the exception-related code -  
we opted for not restricting the analysis only on defect issues.

@@

The Crawler retrieved all issues and comments for these projects.
From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further cleanup (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stacks (regardless its labels).

@@

The ExceptionMiner is a mining tool which can connect to different issue repositories, 
extract issues, mine exception stack traces from them, distill exception stack trace information,
and enable the execution of different analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:

\emph{\textbf{Repository Connectors.}} This component enables the connection 
with issue repositories. In this study two main connectors were created: one which connects to 
GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
 that can traverse the Google Code web interface and extract the project issues. 
Projects meta-data and the issues associated to each project are stored on a relational 
database.

\emph{\textbf{Exception Stack Trace Distiller.}}
This component is based on a combination between a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces mixed with log information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.}.

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analyze the source code of Java Exception classes (when the
bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed all exceptions defined on Android platform (API level 19) - 
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover, it also analyzed the 
exceptions reported on stack traces that were defined on applications and third-party libraries 
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module with the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could refer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, the tool recursively analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source code of reused libraries.
Then, based on this information and using pattern matching 
between the signaler name and the packages, this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were neither defined 
on Android platform, nor on core libraries, nor on applications.
Table~\ref{tab:signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
\label{sec:manual}
In the context of the ExceptionMiner tool, manual inspections were used to:
(i) support on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and (ii)  
identify the type of some exceptions reported on issues, 
that were not be automatically identified by the ExceptionMiner tool
(because they were defined on previous versions of libraries,
apps and Android Plaform). When the exception could not be  
found neither automatically, nor manually (because they were defined on a previous version
of the app or lib) the exception were classified 
in our study as ``Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

\subsection{Replication Package}
All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer.
Specifically we provide: (i) all issues related to Android projects found
in GITHub and Google Code used in this study; (iii) all stack traces extracted
from issues (ii) the results of manual inspection steps; (iv) the tool developed 
in this work to support stack trace extraction and distilling.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guided the search for \emph{bug hazards} on the exception-related code
based on exception stack trace information.

@@

After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.
